\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-04-05 doc/paper.tex
remove "Not yet implemented" remarks from the paper
reformulate calculation of pole masses

The physical (pole) mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices to calculate
the pole masses at the one-loop level, which means finding the
eigenvalues $p$ that solve the following equations:
%
\begin{align}
  &\text{scalars } \phi:  & 0 &= p^2\unity - m^2_{\phi,1L}(p^2), \\
  &\text{fermions } \xi: & 0 &= p\unity - m_{\xi,1L}(p^2),
\end{align}
%
where the one-loop mass matrices are given by
%
\begin{align}
  &\text{scalars } \phi: &
  m^2_{\phi,1L}(p^2) &= m^2_{\phi} - \Sigma_\phi(p^2), \\
  &\text{majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Since the one-loop mass matrices depend on $p$, an iterative procedure
must be used.  \fs provides three different methods with different
precision and execution speed to determine the mass eigenvalues:

\code{LowDiagonalizationPrecision}: This option provides the lowest
precision but is also the fastest one.  Here the one-loop mass matrix
$m_{\phi,1L,\text{low}}$ is calculated exactly once as
%
\begin{align}
  (m_{\phi,1L,\text{low}})_{ij} = (m_{\phi,1L}(p^2 = m_{\phi_i} m_{\phi_j}))_{ij} ,
\end{align}
%
where $m_{\phi_i}$ is the $i$th mass eigenvalue of the tree-level mass
matrix $m_\phi$.  Afterwards $m_{\phi,1L,\text{low}}$ is diagonalized
and the eigenvalues are interpreted as pole masses $m_{\phi_i}^\pole$.

\code{MediumDiagonalizationPrecision} (default): This option provides
calculation with medium precision with a medium execution time.  Here
the one-loop mass matrix $m_{\phi,1L,\text{medium}}$ is calculated $n$
times as
%
\begin{align}
  (m_{\phi,1L,\text{medium}})_{ij}^{(k)} = (m_{\phi,1L}(p^2 = m_{\phi_k}^2))_{ij} ,
  \qquad k = 1,\ldots,n ,
\end{align}
%
where $m_{\phi_k}$ is the $k$th mass eigenvalue of the tree-level mass
matrix $m_\phi$.  Afterwards, each mass matrix
$(m_{\phi,1L,\text{medium}})^{(k)}$ is diagonalized and the $k$th
eigenvalue is interpreted as pole mass $m_{\phi_k}^\pole$.

\code{HighDiagonalizationPrecision}: This option provides
diagonalization with highest precision but has also the highest
execution time.  Here the one-loop mass matrix
$m_{\phi,1L,\text{high}}$ is diagonalized $n$ times as in the case of
\code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
$m_{\phi_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the
diagonalization is repeated, this time using the so calculated pole
masses $m_{\phi_k}^\pole$ for the momentum calculation $p^2 =
(m_{\phi_k}^\pole)^2$, until convergence is reached.

If higher accuracy is required additional routines with higher order
corrections can be added in a very simple way. For example in the MSSM
by default \fs adds two-loop Higgs FORTRAN routines supplied by
P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
$\oatab$ as follows... {\color{red} This is not yet implemented!}

Something similar is done for the NMSSM, but in this case the NMSSM
$\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for
$\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces
are used though it should be understood that these are not complete in
the NMSSM. For other models since the Higgs mass is a very important
measurement and the two loop corrections can be larger than the
current experimental error \cite{Degrassi:2009yq} we recommend that
the leading log two loop corrections are estimated, by generalising
those of the MSSM or NMSSM, as has been done, for example, in the
E$_6$SSM\cite{King:2005jy}.  {\color{red} This is not yet implemented!}

>>>>

The physical (pole) mass spectrum can now be calculated.  \fs uses the
full one-loop self-energies and tree-level mass matrices to calculate
the pole masses at the one-loop level, which means finding the values
$p$ that solve the equation
%
\begin{align}
  0 = \det\left[p^2\unity - m^2_{f,1L}(p^2)\right],
\end{align}
%
where the one-loop mass matrices $m_{f,1L}(p^2)$ are given in terms of
the tree-level mass matrices $m_f$ and the self-energies
$\Sigma_f(p^2)$ as
%
\begin{align}
  &\text{scalars } \phi: &
  m^2_{\phi,1L}(p^2) &= m^2_{\phi} - \Sigma_\phi(p^2), \\
  &\text{majorana fermions } \chi: &
  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
  \Big], \\
  &\text{dirac fermions } \psi: &
  m_{\psi,1L}(p^2) &= m_{\psi}
  - \Sigma_\psi^S(p^2)
  - \Sigma_\psi^R(p^2) m_{\psi}
  - m_{\psi} \Sigma_\psi^L(p^2) .
\end{align}
%
Since the one-loop mass matrices depend on $p$, an iterative procedure
must be used.  \fs provides three different methods with different
precision and execution speed to determine the mass eigenvalues:

\code{LowDiagonalizationPrecision}: This option provides the lowest
precision but is also the fastest one.  Here the one-loop mass matrix
$m_{f,1L}^\text{low}$ is calculated exactly once as
%
\begin{align}
  \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i} m_{f_j}))_{ij} ,
\end{align}
%
where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.

\code{MediumDiagonalizationPrecision} (default): This option provides
calculation with medium precision with a medium execution time.  Here
the one-loop mass matrix $m_{f,1L}^\text{medium}$ is calculated $n$
times as
%
\begin{align}
  (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 = m_{f_k}^2))_{ij} ,
  \qquad k = 1,\ldots,n ,
\end{align}
%
where $m_{f_k}$ is the $k$th mass eigenvalue of the tree-level mass
matrix $m_f$.  Afterwards, each mass matrix
$(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.

\code{HighDiagonalizationPrecision}: This option provides
diagonalization with highest precision, but has also the highest
execution time.  Here the one-loop mass matrix $m_{f,1L}^\text{high}$
is diagonalized $n$ times, as in the case of
\code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
$m_{f_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the diagonalization
is repeated, this time using the calculated pole masses
$m_{f_k}^\pole$ for the momentum calculation $p^2 =
(m_{f_k}^\pole)^2$.  The iteration stops until convergence is reached.

If higher accuracy is required additional routines with higher order
corrections can be added in a very simple way. For example in the MSSM
by default \fs adds two-loop Higgs FORTRAN routines supplied by
P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
$\oatab$ as follows...

Something similar is done for the NMSSM, but in this case the NMSSM
$\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for
$\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces
are used though it should be understood that these are not complete in
the NMSSM. For other models since the Higgs mass is a very important
measurement and the two loop corrections can be larger than the
current experimental error \cite{Degrassi:2009yq} we recommend that
the leading log two loop corrections are estimated, by generalising
those of the MSSM or NMSSM, as has been done, for example, in the
E$_6$SSM\cite{King:2005jy}.
